---
title: "Symbiont_Data"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(devtools)   # install.packages("devtools")
library(steponeR) # install_github("jrcunning/steponeR")
library(tidyverse) # (install.packages("tidyverse")
```

#Importing Raw qPCR Data from Quant Studio Design and Analysis Software 
```{r}
# List Exuma symbiodinium data files
plates <- list.files(path = "Data/qPCR", pattern = ".txt", full.names = TRUE)
plates
# Read in data and calculate target ratios
df <- steponeR(files = plates,
               delim = "\t", 
               target.ratios = "A.A", 
               fluor.norm = list(C = 2.234, D = 1, A = 1, B = 1),
               copy.number = list(C = 20, D = 3, A = 1, B = 1),
               ploidy = list(C = 1, D = 1, A = 1, B = 1),
               extract = list(C = 0.813, D = 0.813, A = 0.813, B = 0.813))
qpcr_stepone <- df$result
```

# Establishing Ratios
```{r}
# adjust CT for flouresence 
qpcr_step_flor <- qpcr_stepone %>%
  mutate(A.CT.mean = A.CT.mean - 1, 
         B.CT.mean = B.CT.mean - 1,
         C.CT.mean = C.CT.mean - 1,
         D.CT.mean = D.CT.mean - 1)

# adjust CT to form ratio of Symbiont divided by Copy # (Currently a Placeholder #)
qpcr_copy_num <- qpcr_step_flor %>%
  mutate(N = 15) %>%
  mutate(A.N = 2^(N - A.CT.mean)/1,
         B.N = 2^(N - B.CT.mean)/1,
         C.N = 2^(N - C.CT.mean)/1,
         D.N = 2^(N - D.CT.mean)/1)

qpcr_prop <- qpcr_copy_num %>%
  mutate(A.N = ifelse(is.na(A.N), 0, A.N),
         B.N = ifelse(is.na(B.N), 0, B.N),
         C.N = ifelse(is.na(C.N), 0, C.N),
         D.N = ifelse(is.na(D.N), 0, D.N)) %>%
  rowwise() %>%
  mutate(total = sum(A.N + B.N + C.N + D.N))%>%
  mutate(A.prop = A.N/total, B.prop = B.N/total, C.prop = C.N/total, D.prop = D.N/total)

qpcr_step <- qpcr_prop %>%
  mutate(dom_sym = case_when(B.prop > C.prop & B.prop > D.prop & B.prop > A.prop ~ "Breviolum",
                             C.prop > B.prop & C.prop > D.prop & C.prop > A.prop ~ "Cladocopium",
                             D.prop > B.prop & D.prop > C.prop & D.prop > A.prop ~ "Durusdinium",
                             A.prop > B.prop & A.prop > C.prop & A.prop > D.prop ~ "Symbiodinium"))
   
```


# Oranizing Raw qPCR Data
```{r}
# Convert NaN CT values to 0  
qpcr <- qpcr_step %>%
 mutate(A.CT.mean = ifelse(is.na(A.CT.mean), 0, A.CT.mean),
         B.CT.mean = ifelse(is.na(B.CT.mean), 0, B.CT.mean),
         C.CT.mean = ifelse(is.na(C.CT.mean), 0, C.CT.mean),
         D.CT.mean = ifelse(is.na(D.CT.mean), 0, D.CT.mean))

# Samples that did not amplify for any symbionts
fails <- qpcr %>%
  filter(A.CT.mean == 0 & B.CT.mean == 0 & C.CT.mean == 0 & D.CT.mean == 0,
         Sample.Name != "PTC",
         Sample.Name != "NTC")

# Samples that need to be rerun due to 1 technical replicate or SD > 1
reruns <- qpcr %>%
  filter(Sample.Name != "PTC", Sample.Name != "NTC") %>%
  filter(A.reps == 1 | B.reps == 1 | C.reps == 1 | D.reps == 1 | A.CT.sd > 1 | B.CT.sd > 1 | C.CT.sd > 1 | D.CT.sd > 1)
```

#Filtering out "bad" data/replicates
```{r}
#Determing which samples have 2 replicates in the data set and did not fail
dupes <- qpcr_step %>%
  filter(Sample.Name != "PTC", Sample.Name != "NTC") %>%
  filter(!(A.CT.mean == 0 & B.CT.mean == 0 & C.CT.mean == 0 & D.CT.mean == 0)) %>%
  group_by(Sample.Name) %>%
  count(vars = "Sample.Name") %>%
  filter(n > 1) 

#Gathering the full data for samples with 2 replicates
exumas_dupes <- qpcr_step %>%
  filter(Sample.Name %in% dupes$Sample.Name) %>%
  arrange(Sample.Name)

#filtering out data with only 1 technical replicate
exumas_dupes_f1 <- exumas_dupes %>%
  group_by(Sample.Name) %>%
  filter(A.reps != 1, B.reps != 1, C.reps != 1, D.reps != 1) %>%
  ungroup()
  
# Average SD  
exumas_dupes_f2 <- exumas_dupes_f1 %>%
  rowwise() %>%
  mutate(avg_sd = mean(c(A.CT.sd, B.CT.sd, C.CT.sd, D.CT.sd), na.rm = TRUE)) %>%
  group_by(Sample.Name) %>%
  dplyr::filter(avg_sd == min(avg_sd)) 

# Raw Data fine on the first run
qpcr_good <- qpcr %>%
  filter(Sample.Name != "PTC", Sample.Name != "NTC") %>%
  filter(!(Sample.Name %in% reruns$Sample.Name))%>%
  filter(!(Sample.Name %in% fails$Sample.Name))


# Cleaned dataset from data that worked the first run and "clean" replicates 
exumas_good <- qpcr_good %>%
  filter(!Sample.Name %in% dupes$Sample.Name) %>%
  bind_rows(exumas_dupes_f2)

need_2_clean <- exumas_good %>%
  group_by(Sample.Name) %>%
  count(vars = "Sample.Name") %>%
  filter(n > 1) 


# Which Samples Were Duplicated and Still Need to be Rerun?
setdiff(exumas_dupes$Sample.Name, exumas_dupes_f2$Sample.Name)

# Samples that need to be rerun
# view(setdiff(reruns$Sample.Name, exumas_dupes_f2$Sample.Name))
```

```{r}
# Import Sample Data 
library(readr)

raw_sample_data <- readxl::read_xlsx("Data/Sample_data/Ex_Sample_Data.xlsx")

#renaming  colomns of sample data to match qpcr data
sample_data <- raw_sample_data %>%
  select(Sample.Name = sample_id, site, genus, species, depth = colony_depth_m, range)

# Joining sample data with qpcr data and removing NA samples
exumas_data <- right_join(sample_data, exumas_good, by = "Sample.Name", all = TRUE)
```

# Distribution of Samples
```{r}
species_count <- exumas_data %>%
  select(Sample.Name, site, genus, depth) %>%
  group_by(genus) %>%
  count(vars = "genus") 

site_count <- exumas_data %>%
  select(Sample.Name, site, species, depth) %>%
  group_by(site) %>%
  count(vars = "site")

depth_metrics<- exumas_data %>%
  group_by(site) %>%
  summarize(
    mean = mean(depth, na.rm = TRUE),
    min = min(depth, na.rm = TRUE),
    max = max(depth, na.rm = TRUE),
    median = median(depth, na.rm = TRUE))


# Visualizing differences in depth
exumas_data %>%
  select(site, depth) %>%
  ggplot(aes(x = site, y = -(depth), fill = site)) +
  geom_boxplot()
  
```

#Symbiont Composition per Sample
```{r}
sym_commun <- exumas_data %>%
  select(Sample.Name, A.prop, B.prop, C.prop, D.prop, depth, site, genus, range) %>%
  mutate(Symbiodinium = case_when(A.prop > 0 ~ A.prop),
         Breviolum = case_when(B.prop > 0 ~ B.prop),
         Cladocopium = case_when(C.prop > 0 ~ C.prop),
         Durusdinium = case_when(D.prop > 0 ~ D.prop)) %>%
  select(Sample.Name, Symbiodinium, Breviolum, Cladocopium, Durusdinium, depth, site, genus) %>%
  gather(key = "sym", value = "value", -Sample.Name, -depth, -site, -genus)

sym_commun_depth <-sym_commun %>%
  filter(depth != "NA") %>%
  mutate(value = ifelse(is.na(value), 0, value))%>%
  mutate(newcol = paste(depth, Sample.Name))

# Symbiont Comp All Samples by Increasing Depth 
# ~5 Samples not represented bc no associated depth recorded
ggplot(sym_commun_depth, aes(x = newcol, y = value, fill = sym)) +
  geom_bar(stat = "identity", position = "fill") +
  scale_fill_manual(values = c('#a6cee3','#1f78b4','#b2df8a','#33a02c'))

# Symbiont Comp by Species and Site 
ggplot(sym_commun, aes(x = Sample.Name, y = value, fill = sym)) +
  geom_bar(stat = "identity", position = "fill") +
  xlab("Sample") +
  ylab("Proportion of Symbionts in Coral") +
  labs(fill = "") +
  facet_grid(genus~site) +
  scale_fill_manual(values = c('#a6cee3','#1f78b4','#b2df8a','#33a02c')) +
  theme_minimal() +
  theme(axis.title.x = element_text(size = 25),
        axis.title.y = element_text(size = 25),
        axis.text.x = element_text(size = 25, color = "transparent", margin = margin(t = -20)),
        axis.text.y = element_text(size = 25, color = "black", margin = margin(l = 30)),
        panel.background = element_rect(color ="black"),
        panel.border = element_rect(color = "black", fill = NA, size = 1.5),
        legend.text = element_text(size = 25, face = "italic"),
        legend.position = "bottom")
```

#Symbiont Composition of Each Sample Grouped by Species
```{r}
# Siderastrea
sid_samples <- sym_commun_depth %>%
  filter(genus == "siderastrea")

sid <- ggplot(sid_samples, aes(x = newcol, y = value, fill = sym)) +
  geom_bar(stat = "identity", position = "fill") +
  facet_wrap(~genus) +
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_manual(values = c('#a6cee3','#1f78b4','#b2df8a','#33a02c')) +
  theme_bw() +
  theme(axis.text.x = element_text(color = "transparent"),
        axis.title.x = element_text(color = "transparent"),
        legend.position = "none")

# Orbicella 
orb_samples <- sym_commun_depth %>%
  filter(genus == "orbicella")

orb <- ggplot(orb_samples, aes(x = newcol, y = value, fill = sym)) +
  geom_bar(stat = "identity", position = "fill") +
  facet_wrap(~genus) +
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_manual(values = c('#a6cee3','#1f78b4','#b2df8a','#33a02c')) +
  theme_bw() +
  theme(axis.text.x = element_text(color = "transparent"),
        axis.title.x = element_text(color = "transparent"),
        legend.position = "none")

# Diploria
dip_samples <- sym_commun_depth %>%
  filter(genus == "diploria")

dip <- ggplot(dip_samples, aes(x = newcol, y = value, fill = sym)) +
  geom_bar(stat = "identity", position = "fill") +
  facet_wrap(~genus) +
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_manual(values = c('#a6cee3','#1f78b4','#b2df8a','#33a02c')) +
  theme_bw() +
  theme(axis.text.x = element_text(color = "transparent"),
        axis.title.x = element_text(color = "transparent"),
        legend.position = "none")

#Pseudodiploria
pse_samples <- sym_commun_depth %>%
  filter(genus == "pseudodiploria")

pse <- ggplot(pse_samples, aes(x = newcol, y = value, fill = sym)) +
  geom_bar(stat = "identity", position = "fill") +
  facet_wrap(~genus) +
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_manual(values = c('#a6cee3','#1f78b4','#b2df8a','#33a02c')) +
  theme_bw() +
  theme(axis.text.x = element_text(color = "transparent"),
        axis.title.x = element_text(color = "transparent"),
        legend.position = "bottom")
```

# Symbiont Composition Each Sample by Species Grid 
```{r}
library(ggpubr) #instal.packages("ggpubr")

ggarrange(sid, orb, dip, pse, 
                    labels = c("  a", "  b","  c","  d" )) +
  theme(legend.position = "bottom")
```

#Symbiont Composition of Each Sample Grouped by Species and Depth Range

```{r}

ranges <- sym_commun %>%
  filter(depth != "NA") 


  mutate(value = ifelse(is.na(value), 0, value))%>%
  mutate(newcol = paste(depth, Sample.Name))
 
  sym_commun_depth <- sym_renamed %>%
  filter(depth != "NA") %>%
  mutate(value = ifelse(is.na(value), 0, value))%>%
  mutate(newcol = paste(depth, Sample.Name))
  
  
  
   filter(range == "shallow") %>%
  select(Sample.Name, genus, range, Symbiodinium, Breviolum, Cladocopium, Durusdinium)
 

```







# Depth of Dominant Symbiont all Data together 
```{r}
dom_sym <- exumas_data %>%
  select(Sample.Name, depth, dom_sym, site) %>%
  filter(depth != "NA") %>%
  mutate(depth_int = cut(depth, breaks = 6)) %>%
  group_by(depth_int, dom_sym, site) %>%
  summarise(ncol = n()) %>%
  mutate(n = sum(ncol), label = paste0("N =", n)) %>%
  mutate(prop = ncol/n)


ggplot(dom_sym, aes(x = as.factor(depth_int), y = ncol, fill = dom_sym)) +
  geom_bar(stat = "identity", position = "fill", size =20) +
  geom_text(aes(x = as.factor(depth_int), y = 1.05, label =label), inherit.aes = FALSE) +
  scale_fill_manual(values = c('#a6cee3','#1f78b4','#b2df8a','#33a02c'))


  
```

# Depth and Dominant Symbiont by Genus
```{r}
dom_sym_species <- exumas_data %>%
  select(Sample.Name, depth, dom_sym, genus) %>%
  filter(depth != "NA") %>%
  mutate(depth_int = cut(depth, breaks = 6)) %>%
  group_by(depth_int, dom_sym, genus) %>%
  summarise(ncol = n()) %>%
  mutate(n = sum(ncol), label = paste0("N =", n)) %>%
  mutate(prop = ncol/n)

ggplot(dom_sym_species, aes(x = as.factor(depth_int), y = ncol, fill = dom_sym)) +
  geom_bar(stat = "identity", position = "fill", size =20) +
  geom_text(aes(x = as.factor(depth_int), y = 1.05, label = label), inherit.aes = FALSE) +
  facet_grid(~genus) +
  scale_fill_manual(values = c('#a6cee3','#1f78b4','#b2df8a','#33a02c'))
```

#Diploria Network Figure 
```{r}
library(igraph) #install.packages("igraph")
library(fields)

net_data_dip <- exumas_data %>%
  select(Sample.Name, genus, 
         Symbiodinium = A.prop, Breviolum = B.prop, 
         Cladocopium = C.prop, Durusdinium = D.prop, depth) %>%
  filter(genus == "diploria", depth != "NA") %>%
  select(Sample.Name, Symbiodinium, Breviolum, Cladocopium, Durusdinium)

edges_dip <- net_data_dip %>%
  gather(key = "Genera", value = "weight", -Sample.Name) %>%
  filter(weight != 0)

depth_dip <- exumas_data %>%
  select(Sample.Name, depth, genus) %>%
  filter(genus == "diploria", depth != "NA") %>%
  select(Sample.Name, depth)

verts_dip <- tibble(Sample.Name = c(net_data_dip$Sample.Name, colnames(net_data_dip)[-1])) %>%
  left_join(depth_dip) %>%
  rename(id = Sample.Name) %>%
  mutate(lab = case_when(grepl("EX", id) ~ "", grepl("um", id) ~ str_sub(id, start = 1, end = 1))) %>%
  select(id, depth, lab)

net_dip <- graph_from_data_frame(edges_dip, directed = FALSE, vertices = verts_dip)

png("Figures/dip_Network.png", width = 200, height = 200, units = "mm", res = 600)

set.seed(6)

plot(net_dip, edge.curved = 0, 
     edge.width = 15 * (E(net_dip)$weight)^0.5, 
     edge.color = "gray65",
     vertex.label = V(net_dip)$lab,
     vertex.label.font = 2,
     vertex.label.cex = 2.5,
     vertex.label.color = "black", 
     vertex.size = ifelse(grepl("EX", names(V(net_dip))), 10, 25),
     vertex.color = ifelse(grepl("EX", names(V(net_dip))), 
                           colorRampPalette(c("white", "black"))(100)[as.numeric(cut(V(net_dip)$depth,breaks = 100))],
                           c('#a6cee3','#1f78b4','#b2df8a','#33a02c')))

image.plot(legend.only = TRUE, 
           zlim = range(V(net_dip)$depth, na.rm = TRUE),
           horizontal = TRUE,
           legend.shrink = 0.6,
           legend.lab = "Depth (m)",
           col = colorRampPalette(c("white", "black"))(100))


dev.off()
```

# Orbicella Network Figure 
```{r}
net_data_orb <- exumas_data %>%
  select(Sample.Name, genus, 
         Symbiodinium = A.prop, Breviolum = B.prop, 
         Cladocopium = C.prop, Durusdinium = D.prop, depth) %>%
  filter(genus == "orbicella", depth != "NA") %>%
  select(Sample.Name, Symbiodinium, Breviolum, Cladocopium, Durusdinium)

edges_orb <- net_data_orb %>%
  gather(key = "Genera", value = "weight", -Sample.Name) %>%
  filter(weight != 0)

depth_orb <- exumas_data %>%
  select(Sample.Name, depth, genus) %>%
  filter(genus == "orbicella", depth != "NA") %>%
  select(Sample.Name, depth)

verts_orb <- tibble(Sample.Name = c(net_data_orb$Sample.Name, colnames(net_data_orb)[-1])) %>%
  left_join(depth_orb) %>%
  rename(id = Sample.Name) %>%
  mutate(lab = case_when(grepl("EX", id) ~ "", grepl("um", id) ~ str_sub(id, start = 1, end = 1))) %>%
  select(id, depth, lab)

net_orb <- graph_from_data_frame(edges_orb, directed = FALSE, vertices = verts_orb)

png("Figures/orb_Network.png", width = 200, height = 200, units = "mm", res = 600)

set.seed(6)

plot(net_orb, edge.curved = 0, 
     edge.width = 15 * (E(net_orb)$weight)^0.5, 
     edge.color = "gray65",
     vertex.label = V(net_orb)$lab,
     vertex.label.font = 2,
     vertex.label.cex = 2.5,
     vertex.label.color = "black", 
     vertex.size = ifelse(grepl("EX", names(V(net_orb))), 10, 25),
     vertex.color = ifelse(grepl("EX", names(V(net_orb))), 
                           colorRampPalette(c("white", "black"))(100)[as.numeric(cut(V(net_orb)$depth,breaks = 100))],
                           c('#a6cee3','#1f78b4','#b2df8a','#33a02c')))

image.plot(legend.only = TRUE, 
           zlim = range(V(net_orb)$depth, na.rm = TRUE),
           horizontal = TRUE,
           legend.shrink = 0.6,
           legend.lab = "Depth (m)",
           col = colorRampPalette(c("white", "black"))(100))


dev.off()
```

#Pseudodiploria Network
```{r}

net_data_pse <- exumas_data %>%
  select(Sample.Name, genus, 
         Symbiodinium = A.prop, Breviolum = B.prop, 
         Cladocopium = C.prop, Durusdinium = D.prop, depth) %>%
  filter(genus == "pseudodiploria", depth != "NA") %>%
  select(Sample.Name, Symbiodinium, Breviolum, Cladocopium, Durusdinium)

edges_pse <- net_data_pse %>%
  gather(key = "Genera", value = "weight", -Sample.Name) %>%
  filter(weight != 0)

depth_pse <- exumas_data %>%
  select(Sample.Name, depth, genus) %>%
  filter(genus == "pseudodiploria", depth != "NA") %>%
  select(Sample.Name, depth)

verts_pse <- tibble(Sample.Name = c(net_data_pse$Sample.Name, colnames(net_data_pse)[-1])) %>%
  left_join(depth_pse) %>%
  rename(id = Sample.Name) %>%
  mutate(lab = case_when(grepl("EX", id) ~ "", grepl("um", id) ~ str_sub(id, start = 1, end = 1))) %>%
  select(id, depth, lab)

net_pse <- graph_from_data_frame(edges_pse, directed = FALSE, vertices = verts_pse)

png("Figures/pse_Network.png", width = 200, height = 200, units = "mm", res = 600)

set.seed(6)

plot(net_pse, edge.curved = 0, 
     edge.width = 15 * (E(net_pse)$weight)^0.5, 
     edge.color = "gray65",
     vertex.label = V(net_pse)$lab,
     vertex.label.font = 2,
     vertex.label.cex = 2.5,
     vertex.label.color = "black", 
     vertex.size = ifelse(grepl("EX", names(V(net_pse))), 10, 25),
     vertex.color = ifelse(grepl("EX", names(V(net_pse))), 
                           colorRampPalette(c("white", "black"))(100)[as.numeric(cut(V(net_pse)$depth,breaks = 100))],
                           c('#a6cee3','#1f78b4','#b2df8a','#33a02c')))

image.plot(legend.only = TRUE, 
           zlim = range(V(net_pse)$depth, na.rm = TRUE),
           horizontal = TRUE,
           legend.shrink = 0.6,
           legend.lab = "Depth (m)",
           col = colorRampPalette(c("white", "black"))(100))


dev.off()
```

#Siderastrea Network 
```{r}

net_data_sid <- exumas_data %>%
  select(Sample.Name, genus, 
         Symbiodinium = A.prop, Breviolum = B.prop, 
         Cladocopium = C.prop, Durusdinium = D.prop, depth) %>%
  filter(genus == "siderastrea", depth != "NA") %>%
  select(Sample.Name, Symbiodinium, Breviolum, Cladocopium, Durusdinium)

edges_sid <- net_data_sid %>%
  gather(key = "Genera", value = "weight", -Sample.Name) %>%
  filter(weight != 0)

depth_sid <- exumas_data %>%
  select(Sample.Name, depth, genus) %>%
  filter(genus == "siderastrea", depth != "NA") %>%
  select(Sample.Name, depth)

verts_sid <- tibble(Sample.Name = c(net_data_sid$Sample.Name, colnames(net_data_sid)[-1])) %>%
  left_join(depth_sid) %>%
  rename(id = Sample.Name) %>%
  mutate(lab = case_when(grepl("EX", id) ~ "", grepl("um", id) ~ str_sub(id, start = 1, end = 1))) %>%
  select(id, depth, lab)

net_sid <- graph_from_data_frame(edges_sid, directed = FALSE, vertices = verts_sid)

png("Figures/sid_Network.png", width = 200, height = 200, units = "mm", res = 600)

set.seed(10)

plot(net_sid, edge.curved = 0, 
     edge.width = 15 * (E(net_sid)$weight)^0.5, 
     edge.color = "gray65",
     vertex.label = V(net_sid)$lab,
     vertex.label.font = 2,
     vertex.label.cex = 2.5,
     vertex.label.color = "black", 
     vertex.size = ifelse(grepl("EX", names(V(net_sid))), 10, 25),
     vertex.color = ifelse(grepl("EX", names(V(net_sid))), 
                           colorRampPalette(c("white", "black"))(100)[as.numeric(cut(V(net_sid)$depth,breaks = 100))],
                           c('#a6cee3','#1f78b4','#b2df8a','#33a02c')))

image.plot(legend.only = TRUE, 
           zlim = range(V(net_sid)$depth, na.rm = TRUE),
           horizontal = TRUE,
           legend.shrink = 0.6,
           legend.lab = "Depth (m)",
           col = colorRampPalette(c("white", "black"))(100))


dev.off()
```
