---
title: "marine_research_trip_samples"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(devtools)   # install.packages("devtools")
library(steponeR) # install_github("jrcunning/steponeR")
library(tidyverse) # (install.packages("tidyverse")
```


```{r}
marine_trip_data <- readxl::read_xlsx("sample_database(1).xlsx")

mtsites <- marine_trip_data %>%
  filter(site == "parrotfish_reef"| site == "leafy_cay_1"| site == "leafy_cay_2") %>%
  select(sample_id, site, genus, species) %>%
  filter(genus == "orbicella"| genus == "siderastrea")

speciescount <- mtsites %>%
  count()
```


```{r}
#Importing Raw qPCR Data from Quant Studio Design and Analysis Software 
```{r}
# List Exuma symbiodinium data files
marine_trip_data <- list.files(path = "Data/qPCR", pattern = ".txt", full.names = TRUE)
marine_trip_data
# Read in data and calculate target ratios
df <- steponeR(files = marine_trip_data,
               delim = "\t", 
               target.ratios = "A.A", 
               fluor.norm = list(C = 2.234, D = 1, A = 1, B = 1),
               copy.number = list(C = 20, D = 3, A = 1, B = 1),
               ploidy = list(C = 1, D = 1, A = 1, B = 1),
               extract = list(C = 0.813, D = 0.813, A = 0.813, B = 0.813))
qpcr_marinetrip <- df$result %>%
  filter(File.Name == "marinetrip.txt") %>%
  rename(sample_id = Sample.Name)
```

# Establishing Ratios
```{r}
# adjust CT for flouresence 
qpcr_flor <- qpcr_marinetrip %>%
  mutate(A.CT.mean = A.CT.mean - 1, 
         B.CT.mean = B.CT.mean - 1,
         C.CT.mean = C.CT.mean - 1,
         D.CT.mean = D.CT.mean - 1)

# adjust CT to form ratio of Symbiont divided by Copy # (Currently a Placeholder #)
qpcr_copynumber <- qpcr_flor %>%
  mutate(N = 15) %>%
  mutate(A.N = 2^(N - A.CT.mean)/1,
         B.N = 2^(N - B.CT.mean)/1,
         C.N = 2^(N - C.CT.mean)/1,
         D.N = 2^(N - D.CT.mean)/1)

qpcr_proportions <- qpcr_copynumber %>%
  mutate(A.N = ifelse(is.na(A.N), 0, A.N),
         B.N = ifelse(is.na(B.N), 0, B.N),
         C.N = ifelse(is.na(C.N), 0, C.N),
         D.N = ifelse(is.na(D.N), 0, D.N)) %>%
  rowwise() %>%
  mutate(total = sum(A.N + B.N + C.N + D.N))%>%
  mutate(A.prop = A.N/total, B.prop = B.N/total, C.prop = C.N/total, D.prop = D.N/total)

qpcr_dominant_sym <- qpcr_proportions %>%
  mutate(dom_sym = case_when(B.prop > C.prop & B.prop > D.prop & B.prop > A.prop ~ "Breviolum",
                             C.prop > B.prop & C.prop > D.prop & C.prop > A.prop ~ "Cladocopium",
                             D.prop > B.prop & D.prop > C.prop & D.prop > A.prop ~ "Durusdinium",
                             A.prop > B.prop & A.prop > C.prop & A.prop > D.prop ~ "Symbiodinium"))
   
```

```{r}

# Joining sample data with qpcr data and removing NA samples
marine_data <- right_join(mtsites, qpcr_dominant_sym, by = "sample_id", all = TRUE)%>%
  select(sample_id, site, genus, species, A.prop, B.prop, C.prop, D.prop, dom_sym)

single_sym <- marine_data %>%
  filter(A.prop == 1| B.prop == 1| C.prop == 1| D.prop == 1)
```

